---
title: "Project 02 - Model Exploration_JAL"
author: "John Leraas"
date: "11/24/2020"
output: html_document
---

```{r}
install.packages("car")
install.packages("ResourceSelection")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(faraway)
library(multcomp)
library(ROCR)
library(car) #used for GVIF (to address categorical variables) https://cran.r-project.org/web/packages/car/car.pdf
library(ResourceSelection) #Used for Hosmer Leschow Test https://www.rdocumentation.org/packages/ResourceSelection/versions/0.3-5/topics/hoslem.test
library(leaps) #For All Regression
```


## Import Data & Clean

- Load Data
- Drop unwanted columns
- Remove rows with "N/A" (only one exists after unwanted columns removed)


```{r}

# read csv file, and remove column 1 and 16, which is index column and ID
nass <- read.csv("nassCDS.csv", header = T)

lengths(nass)

# drop column, X, caseid, weight, injSeverity
nass=subset(nass, select = -c(X,caseid, weight, injSeverity))
lengths(nass)

# check for rows containing NA
which(is.na(nass), arr.ind=TRUE)

# remove rows containing NA
nass <- nass[complete.cases(nass), ]
lengths(nass)

```

## Convert categorical variables to factors

- Ensure categorical variables are factored
- Change "Dead" from ('alive', 'dead') to (0, 1)

```{r}
#convert categorical to factor
nass$dead<-factor(nass$dead) 
nass$dvcat<-factor(nass$dvcat)
nass$airbag<-factor(nass$airbag)
nass$seatbelt<-factor(nass$seatbelt)
nass$sex<-factor(nass$sex)
nass$abcat<-factor(nass$abcat)
nass$occRole<-factor(nass$occRole)
nass$deploy<-factor(nass$deploy)
nass$abcat<-factor(nass$abcat)

nass$dead_num <- nass$dead
levels(nass$dead_num) <- c(0,1)

#remove column of dead
nass=subset(nass, select = -c(dead))
head(nass, 5)
```

## Model Selection
Define full and null models - used in Forward Selection, Backward Elimination, and Stepwise Regression
```{r}
fullmod <- glm(dead_num ~. , data = nass, family = "binomial" ,maxit = 100)
Nullmod <- glm(dead_num~1, data = nass, family = "binomial")
```


All Regression
```{r}
##perform all possible regressions (1st order)
allreg <- regsubsets(dead_num ~., data=nass, nbest=5)

##create a "data frame" that stores the predictors in the various models considered as well as their various criteria
best <- as.data.frame(summary(allreg)$outmat)
best$p <- as.numeric(substr(rownames(best),1,1))+1
best$r2 <- summary(allreg)$rsq
best$adjr2 <- summary(allreg)$adjr2
best$mse <- (summary(allreg)$rss)/(dim(nass)[1]-best$p)
best$cp <- summary(allreg)$cp
best$bic <- summary(allreg)$bic
best

##sort by various criteria
#best[order(best$r2),]
#best[order(best$adjr2),]
best[order(best$mse),]
#best[order(best$cp),]
#best[order(best$bic),]

```


Forward Selection
```{r}
step(Nullmod, scope=list(lower=Nullmod, upper=fullmod), direction="forward")
```

Backward Elimination
```{r}
step(fullmod, scope=list(lower=Nullmod, upper=fullmod), direction="backward")
```

Stepwise Regression - Null Model
```{r}
step(Nullmod, scope=list(lower=Nullmod, upper=fullmod), direction="both")
```


Stepwise Regression - Full Model
Note: I just tried this to see if it resulted in a different model
```{r}
step(fullmod, scope=list(lower=Nullmod, upper=fullmod), direction="both")
```
## Model Comments
Forward Selection, Backward Elimination, and Stepwise Regression each yielded the same model. Additionally Stepwise regression was tried starting with both the Full model and Null model and came to the same result.

Choose the following variables in the model we choose based upon (i) exploratory analysis, and (ii) intuition of important factors (speed of accident, seatbelts, etc).
- dvcat (speed)
- seatbelt
- frontal
- sex
- ageOFocc
- abcat (airbag deployment)
- occRole

## Chosen Model
```{r}
#Test model:
#model_picked <- glm(formula = dead_num ~ dvcat + seatbelt, family = "binomial", data = nass, maxit = 100)

model_picked <- glm(formula = dead_num ~ dvcat + seatbelt + frontal + sex + ageOFocc + 
    abcat + occRole, family = "binomial", data = nass, maxit = 100)

summary(model_picked)
```
## Evaluate Coefficients (Wald Test)
Perform Wald Test on:
dvcat10-24
dvcat25-39
abcatunavail

Wald test: Z = B_j - 0 / se(B_j)
Normal Distribution

dvcat10-24
```{r}
# Wald Test
z_10_24 <- coef(summary(model_picked))["dvcat10-24", "Estimate"] / coef(summary(model_picked))["dvcat10-24", "Std. Error"]
z_10_24
qnorm(0.975)
```
Result: Test Statistic is less extreme than the critical value, so we can omit dvcat10-24


dvcat25-39
```{r}
z_10_24 <- coef(summary(model_picked))["dvcat25-39", "Estimate"] / coef(summary(model_picked))["dvcat25-39", "Std. Error"]
z_10_24
qnorm(0.975)
```
Result: Test Statistic is more extreme than the critical value so we cannot omit dvcat25-39

abcatunavail
```{r}
z_10_24 <- coef(summary(model_picked))["abcatunavail", "Estimate"] / coef(summary(model_picked))["abcatunavail", "Std. Error"]
z_10_24
qnorm(0.975)
```
Result: Test Statistic is less extreme than the critical value, so we can omit abcatunavail

## [Remove Unnecessary Predictors ???????]
Remove
- dvcat10-24
- abcatunavail


## GVIF Function (Using Car Package)

```{r}
faraway::vif(model_picked)
car::vif(model_picked)
```

## Delta G ^ 2
11 coefficients
```{r}
1-pchisq(Nullmod$deviance-model_picked$deviance, 11)
```
Result: Model is useful

## Outliers??

```{r}
COOKS<-cooks.distance(model_picked)
COOKS[COOKS>qf(0.5,P,N-P)]

```

```{r}
P <- 11
N <- 26216
DFFITS<-dffits(model_picked)
DFFITS[abs(DFFITS)>2*sqrt(P/N)]

DFBETAS<-dfbetas(model_picked)
DFBETAS[abs(DFBETAS)>2/sqrt(N)]

```



## Hosmer-Lemeshow Goodness of Fit Test

NOTE: I tried running the following, but not useful for factors (wanted to temporarily document the dead end)

https://www.rdocumentation.org/packages/ResourceSelection/versions/0.3-5/topics/hoslem.test
```{r}
#lengths(model_picked)
hl <- hoslem.test(nass$dead_num, predict(model_picked,data=nass, type="response"), g=10)
#lengths(nass)
```


## Validation

```{r}
#creating training and testing dataset
set.seed(139)
sample<-sample.int(nrow(nass), floor(.50*nrow(nass)), replace = F)
train<-nass[sample, ]
test<-nass[-sample, ]


#Chosen Model
#Test Model:
#model_test <- glm(formula = dead_num ~ dvcat, family = "binomial", data = train, maxit = 100)

model_test <- glm(formula = dead_num ~ dvcat + seatbelt + frontal + sex + ageOFocc + 
    abcat + occRole, family = "binomial", data = train, maxit = 100)

#Null Model
model_null <- glm(dead_num~1, data=train, "binomial", maxit = 100)
summary(model_null)

#ROC
preds<-predict(model_test,newdata=test, type="response")

rates<-prediction(preds, test$dead_num)

roc_result<-performance(rates,measure="tpr", x.measure="fpr")

plot(roc_result, main="ROC Curve for Exploratory Model")
lines(x = c(0,1), y = c(0,1), col="red")


```

## AUC

```{r}
##compute the AUC
aucvalues <- performance(rates, measure = "auc")
auc <- aucvalues@y.values[[1]]
auc
```

## Confusion Matrix

```{r}
table(test$dead_num, preds>0.1)
table(test$dead_num, preds>0.25)
table(test$dead_num, preds>0.5)
table(test$dead_num, preds>0.75)
```


