---
title: "Project 2 - Car Accident Fatalities"
author: "John, Daniel, Jing, Michael"
date: "11/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(faraway) #used for VIF
library(multcomp)
library(ROCR)
library(car) #used for GVIF (to address categorical variables) https://cran.r-project.org/web/packages/car/car.pdf
library(leaps) #For All Regression
```

# Data Cleaning

## Import Data & Clean

- Load Data
- Drop unwanted columns
- Remove rows with "N/A" (only one exists after unwanted columns removed)

Unwanted Columns:
- X: just a number with no corresponding contextual information
- caseid: just an identifier, not contextually relevant
- weight: used in weighting of observations, not used in our model selection process
- injury severity: will assume that this is not known when presence of fatality is unknown

```{r}

# read csv file
nass <- read.csv("nassCDS.csv", header = T)
lengths(nass)

# drop column, X, caseid, weight, injSeverity
nass=subset(nass, select = -c(X,caseid, weight, injSeverity))
lengths(nass)

# check for rows containing NA
which(is.na(nass), arr.ind=TRUE)

# remove rows containing NA
nass <- nass[complete.cases(nass), ]
lengths(nass)

```

## Convert categorical variables to factors

- Ensure categorical variables are factored
- Change "Dead" from ('alive', 'dead') to (0, 1)

```{r}
#convert categorical to factor
nass$dead<-factor(nass$dead) 
nass$dvcat<-factor(nass$dvcat)
nass$airbag<-factor(nass$airbag)
nass$seatbelt<-factor(nass$seatbelt)
nass$sex<-factor(nass$sex)
nass$abcat<-factor(nass$abcat)
nass$occRole<-factor(nass$occRole)
nass$deploy<-factor(nass$deploy)
nass$abcat<-factor(nass$abcat)

nass$dead_num <- nass$dead
levels(nass$dead_num) <- c(0,1)

#remove column of dead
nass=subset(nass, select = -c(dead))
head(nass, 5)
```



# Exploratory Data Analysis
[ADD CODE / ANALYSIS HERE]
**************************
**************************
**************************
[ADD CODE / ANALYSIS HERE]


# Model Selection - Exploration

## Model Selection Exploration - Full Data Set
Define full and null models - used in Forward Selection, Backward Elimination, and Stepwise Regression
```{r}
fullmod <- glm(dead_num ~. , data = nass, family = "binomial" ,maxit = 100)
Nullmod <- glm(dead_num~1, data = nass, family = "binomial")
```

All Regression
```{r}
##perform all possible regressions (1st order)
allreg <- regsubsets(dead_num ~., data=nass, nbest=5)

##create a "data frame" that stores the predictors in the various models considered as well as their various criteria
best <- as.data.frame(summary(allreg)$outmat)
best$p <- as.numeric(substr(rownames(best),1,1))+1
best$r2 <- summary(allreg)$rsq
best$adjr2 <- summary(allreg)$adjr2
best$mse <- (summary(allreg)$rss)/(dim(nass)[1]-best$p)
best$cp <- summary(allreg)$cp
best$bic <- summary(allreg)$bic
best

##sort by various criteria
#best[order(best$r2),]
#best[order(best$adjr2),]
best[order(best$mse),]
#best[order(best$cp),]
#best[order(best$bic),]

```


Backward Elimination
```{r}
step(fullmod, scope=list(lower=Nullmod, upper=fullmod), direction="backward")
```

Stepwise Regression
```{r}
step(Nullmod, scope=list(lower=Nullmod, upper=fullmod), direction="both")
```

Forward Selection
```{r}
# forward selection
step(Nullmod, scope=list(lower=Nullmod, upper=fullmod), direction="forward")
```

### Model Comments:
Forward Selection, Backward Elimination, and Stepwise Regression each yielded the same model. 

Choose the following variables in the model we choose based upon (i) exploratory analysis, and (ii) intuition of important factors (speed of accident, seatbelts, etc. - each seem reasonable).
- dvcat (speed)
- seatbelt
- frontal
- sex
- ageOFocc
- abcat (airbag deployment)
- occRole


Exploratory Model
```{r}
model_picked <- glm(formula = dead_num ~ dvcat + seatbelt + frontal + sex + ageOFocc + 
    abcat + occRole, family = "binomial", data = nass, maxit = 100)

summary(model_picked)
```
Wald Test on individual predictors:
Wald test: Z = B_j - 0 / se(B_j)
Normal Distribution
Note: Can just compare z-value from summary table to qnorm(0.975)
```{r}
qnorm(0.975)
```
Can remove:
-dvcat10-24 (we chose to combine with the reference class, resulting in speed <25km/hr)
-abcatunavail (we chose to not combine with reference class given different contextual interpretations)


Generalized VIF, Delta G^2
```{r}
faraway::vif(model_picked)
car::vif(model_picked)
1-pchisq(Nullmod$deviance-model_picked$deviance, 11)
```

GVIF^(1/2*DF) - low values indicate that multicollinearity is not present
Delta G^2 - indicates that the model is useful (relative to the intercept only model)


## Model Formalization & Rationale

Split Data into Training & Test Set
```{r}
#creating training and testing dataset
set.seed(40)
sample<-sample.int(nrow(nass), floor(.50*nrow(nass)), replace = F)
train<-nass[sample, ]
test<-nass[-sample, ]
```

Automatic Search Processes
```{r}
#Define Full Model and Null Model
fullmod.train <- glm(dead_num ~. , data = train, family = "binomial" ,maxit = 100)
Nullmod.train <- glm(dead_num~1, data = train, family = "binomial")

# Backward Elimination
step(fullmod.train, scope=list(lower=Nullmod.train, upper=fullmod.train), direction="backward")

# Stepwise Regression
step(Nullmod.train, scope=list(lower=Nullmod.train, upper=fullmod.train), direction="both")

# Forward Selection
step(Nullmod.train, scope=list(lower=Nullmod.train, upper=fullmod.train), direction="forward")

```
Note: Each of the processes results in the same model


```{r}
qnorm(0.975) #Critical value for Wald Test

model_final <- glm(formula = dead_num ~ dvcat + seatbelt + frontal + sex + ageOFocc + 
    abcat + occRole, family = "binomial", data = train, maxit = 100)

summary(model_final)
```
Note: Consistent with results from model based on full data set, we can combine dvcat10-24 with the reference class. In addition, it intuitively makes sense that differentiating speeds of accidents that are relatively slow may have little additional predictive power in identifying fatilities.


### Combine appropriate accident speeds 
Combine dvcat 10-24 with reference class. Call 1-25

```{r}

# Full Data Set - combine categories
nass$New_dvcat=nass$dvcat
levels(nass$New_dvcat)<-c(levels(nass$New_dvcat),"1-24")
nass$New_dvcat[nass$dvcat=="1-9km/h"]<-"1-24"
nass$New_dvcat[nass$dvcat=="10-24"]<-"1-24"
nass$New_dvcat <- droplevels(nass$New_dvcat)
levels(nass$New_dvcat)
nass$New_dvcat<- relevel(nass$New_dvcat, ref = "1-24")
levels(nass$New_dvcat)

```

### Refit model based upon new categorization
```{r}
#creating training and testing dataset
set.seed(40) #Note - same sample seed is used for consistency (rather than combining categories for the Test Data set)
sample<-sample.int(nrow(nass), floor(.50*nrow(nass)), replace = F)
train<-nass[sample, ]
test<-nass[-sample, ]

model_final <- glm(formula = dead_num ~ New_dvcat + seatbelt + frontal + sex + ageOFocc + 
    abcat + occRole, family = "binomial", data = train, maxit = 100)

summary(model_final)

```
### Model Validation
ROC Curve
```{r}
#ROC
preds<-predict(model_final,newdata=test, type="response")
rates<-prediction(preds, test$dead_num)
roc_result<-performance(rates,measure="tpr", x.measure="fpr")

plot(roc_result, main="Model ROC Curve")
lines(x = c(0,1), y = c(0,1), col="red")
```

AUC
```{r}
##compute the AUC
aucvalues <- performance(rates, measure = "auc")
auc <- aucvalues@y.values[[1]]
auc
```

Generalized VIF, Delta G^2
```{r}
faraway::vif(model_final)
car::vif(model_final)
1-pchisq(Nullmod.train$deviance-model_final$deviance, 11)
```

GVIF^(1/2*DF) - low values indicate that multicollinearity is not present
Delta G^2 - indicates that the model is useful (relative to the intercept only model)


TO DO:
[CONFUSION MATRICES]
